---
title: 25-35 Study Hours a Week? the Life of a Statistics Major
author: William Medwid
date: '2020-08-26'
slug: 25-35-study-hours-a-week-the-life-of-a-statistics-major
categories:
  - R
tags:
  - Academics
  - Personally Collected Data
subtitle: ''
summary: ''
authors: []
lastmod: '2020-08-26T19:03:32-07:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r libraries, echo=FALSE}
library(tidyverse)
library(stringr)
library(DT)
library(lubridate)
library(modelr)
```

Not all college slogans are the real inspirational type. Cal Poly happens to have a funnily direct one: Study 25-35 hours per week. At least at first while touring the school, it seems to be everywhere from posters on the wall to pins and magnets given out at orientation. 2 hours per week per unit, outside lecture, sounds like it'd add up to a whole lot. So, I've got to ask, does it match up with what students actually do?

As a bit of a spreadsheet nerd, I've found myself in a great position to dive into this question and see if I've at least lived up to the slogan myself. Ever since my second year of high school, consistently recording my time spent on academics in a spreadsheet has been an integral part of my work flow. Each assignment is added as a new row once it's assigned, and I'll do my best to describe it and estimate how long it'll take to complete. This makes a living to-do list where I can punch in and out, recording the time in helper columns. Once the assignment is done and I've recorded every minute spent on it, it's filtered out of sight, clearing the list.

Now, let's take a look at the raw data, as that's often the best way to understand what we're working with. This includes every assignment and class over my first year of college. Feel free to explore the data with the interactive table below, or just carry on for the analysis.

```{r data}
# Note: This data is slightly pre-cleaned so only relevant columns and classes are there, and to avoid excess personal information.

all_quarters <- c("Summer 2019", "Fall 2019", "Winter 2020", "Spring 2020", "Summer 2020")
online_quarters <- c("Spring 2020", "Summer 2020")
units <- c(8,14,16,16,8)
#Best to define these here, they'll matter later.

study_time <- read.csv("collegeFirstYear.csv") %>% 
  mutate(Completed = mdy(Completed))%>% 
  mutate(week_completed = Completed %>% floor_date(unit="week")) %>% 
  mutate(Quarter = factor(Quarter, levels = all_quarters)) %>% 
  mutate(quarter_units = factor(Quarter,
                             levels=all_quarters,
                             labels=units)) %>% 
  filter(Class != "GameDevClub")

DT::datatable(study_time, rownames = FALSE)
```
The three columns Type, Tasks_Predicted, and Units are meant to read sort of like a sentence, with Title sometimes supplementing. For example, the first entry could be interpreted as "I read 14 pages of the Communist Manifesto." A fitting start to my college experience.

The final two columns are for invalidating failures, wherein I didn't complete the task, and imprecise times, where the estimated or real times are approximately correct but not to the precise standards seen elsewhere.

```{r}

workloads <- study_time %>% 
  group_by(week_completed) %>% 
  filter("Class" %in% Type | Quarter %in% online_quarters) %>% #Only counting weeks where there were classes or the quarter was online.
  summarize(week_workload = sum(Min_Real, na.rm=TRUE)/60, Quarter=Quarter, quarter_units = quarter_units) %>% 
  ungroup() %>% 
  distinct(week_completed, .keep_all = TRUE)

workloads %>% 
  ggplot(aes(x=week_completed, y=week_workload, color=Quarter)) +
  geom_point() +
  geom_smooth() +
  xlab("Week") +
  ylab("Hours of Academics Completed") +
  ggtitle("Academic workload over time (including lectures)")
```



```{r}
#Mean is the statistic of choice because large projects can take a long time and end up with work from one week being counted in another.
conf_interval <- t.test(workloads$week_workload, mu=25, alternative = "less")

class_time_total <- study_time %>% 
  filter(Type=="Class") %>% pull(Min_Real) %>% sum(na.rm=TRUE)
study_time_total <- study_time %>% 
  filter(Type!="Class") %>% pull(Min_Real) %>% sum(na.rm=TRUE)
percent_in_class <- (class_time_total / (class_time_total + study_time_total))*100 %>% round(digits=1)
```

```{r include=FALSE}
# bootstrapped for median
# conf_interval <- workloads %>%
#   modelr::bootstrap(count(workloads)$n) %>%
#   mutate(median = map_dbl(strap, ~median(data.frame(.x)$week_workload, na.rm = TRUE))) %>%
#   ungroup() %>%
#   summarize(median_low = quantile(median,0.025), median_high = quantile(median,0.975))
```

Alas, I've only spent a mean of `r conf_interval$estimate` hours per week, a good ways from 25, let alone 35. And that's with classes included, which are `r percent_in_class`% of that time. Apart from the intensive 4-week 8-unit 2019 summer session, there's very little case to be made that I ended up having a typical 25-35 hour week, other than to consider the time spent traveling between classes, which, though quite time consuming, is hardly studying.

```{r}
conf_interval_per_unit <- t.test(workloads$week_workload/units, mu=2, alternative = "less")
```

Another more charitable way to reconcile my schedule with the ideal would be to look at the per-unit time spent. It turns out that's a more positive `r conf_interval_per_unit$estimate` hours, or the equivalent of `r conf_interval_per_unit$estimate*16` hours for a 16-unit schedule. This is still including classes but feels almost like victory nonetheless.

There's probably no need to feel lazy or anything about this - after all, if I can succeed in classes in less thann the expected time, that's probably more an indictment on the classes than on myself. So, let's see, which classes were the ones that never required super rigorous homework?

```{r}
workloads_by_class <- study_time %>% 
  filter(Class != "General") %>% # Removing non class-specific academic imperatives
  group_by(week_completed) %>% 
  filter("Class" %in% Type | Quarter %in% online_quarters) %>% #Only counting weeks where there were classes or the quarter was online.
  ungroup() %>% 
  filter(Type != "Class") %>% 
  group_by(Class,week_completed) %>% 
  summarize(week_workload = sum(Min_Real, na.rm=TRUE)/60, Quarter=Quarter, quarter_units = quarter_units) %>% 
  ungroup() %>% 
  distinct(week_completed, Class, .keep_all = TRUE)

workloads_by_class %>% 
  ggplot(aes(y=Class,x=week_workload)) +
  geom_boxplot(fill="orangered") +
  xlab("Weekly Workload (Excluding Classes)") +
  facet_grid(rows= vars(Quarter), scales = "free_y") +
  geom_vline(xintercept =2, color="darkgreen", linetype = "dashed")+
  scale_x_continuous(breaks=c(0,1,2,5,10,15))
```

The chart shows only three classes below the to hour line, which match my memories of the less rigorous yet still valuable courses. Note that of the outlier groups may be exaggerated in their extremity because a project done slowly over multiple weeks only counts on the week when it's completed, unless I've broken the project into multiple component tasks. 

Feel free to contact me at the home page if you're an architecture major wondering if I've left a factor of 10 off of all of these counts, or someone working a 40 hour work week, having a stroke looking at this.

